---
title: "class15"
author: "Zhuohang Wu"
date: "February 26, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
counts <- read.csv("data/airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <-  read.csv("data/airway_metadata.csv", stringsAsFactors = FALSE)
```

```{r}
head(counts)
```

```{r}
head(metadata)
```


```{r}
control.inds <- metadata$dex=="control"
control.ids <- metadata$id[control.inds]
head(counts[,control.ids])
```

```{r}
control.mean <- rowSums(counts[,control.ids])/ncol(counts[,control.ids])
head(control.mean)
names(control.mean) <- counts$ensgene
```

```{r}
drug.inds <- metadata$dex=="treated"
drug.ids <- metadata$id[drug.inds]
head(counts[,drug.ids])
```

```{r}
drug.mean <- rowSums(counts[,drug.ids])/ncol(counts[,drug.ids])
head(drug.mean)
names(drug.mean) <- counts$ensgene
```

```{r}
mean.counts <- data.frame(control.mean,drug.mean)
head(mean.counts)
```

```{r}
plot(control.mean,drug.mean,xlab = "control",ylab = "drug",main = "Treated vs. Control")
```


```{r}
plot(control.mean,drug.mean,log="xy",xlab = "control",ylab = "drug",main = "Treated vs. Control")
```

We can find candidate differentially expressed genes by looking for genes with a large change between control and dex-treated samples. We usually look at the log2 of the fold change, because this has better mathematical properties.

Here we calculate log2foldchange, add it to our meancounts data.frame and inspect the results either with the head() or the View() function for example.

```{r}
mean.counts$log2fc <- log2(mean.counts[,"drug.mean"]/mean.counts[,"control.mean"])
head(mean.counts)
```

```{r}
t.test(control.mean,drug.mean)
```

```{r}
zero.vals <- which(mean.counts[,1:2]==0, arr.ind=TRUE)
```


```{r}
to.rm <- unique(zero.vals[,1])
mycounts <- mean.counts[-to.rm,]
head(mycounts)
```

```{r}
up.ind <- mycounts$log2fc > 2
down.ind <- mycounts$log2fc < (-2)
```

```{r}
head(mycounts[up.ind,])
```

```{r}
head(mycounts[down.ind,])
```


```{r}
sum(up.ind)
sum(down.ind)
```

```{r}
anno <- read.csv("data/annotables_grch38.csv")
head(anno)
```


```{r}
head(mycounts)
```

```{r}
#head(mycounts)
my.count.annotated <- merge(x=mycounts, y= anno, by.x = "row.names", by.y = "ensgene")
head(my.count.annotated)
```

```{r}
library(DESeq2)
```

```{r}
citation("DESeq2")
```

```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds
```

```{r}
dds <- DESeq(dds)
```

```{r}
res <- results(dds)
res
```

```{r}
summary(res)
```

```{r}
res01 <- results(dds, alpha=0.01)
summary(res01)
```


```{r}
plot(res$log2FoldChange, -log(res$padj))
```

```{r}
resSig01 <- subset(as.data.frame(res), padj < 0.01)
nrow(resSig01)
```

```{r}
#resSig01 <- subset(as.data.frame(res), padj < 0.01)
#nrow(resSig01)
ord <- order( resSig01$padj )
View(res01[ord,])
head(resSig01[ord,])
```

```{r}
#resSig01$
head(resSig01[ord,])
```



```{r}
head(anno)
```

```{r}
resSig01<-merge(x=resSig01[ord,], y=anno, by.x="row.names", by.y="ensgene")
head(resSig01)
```


```{r}
write.csv(resSig01[ord,], "signif01_results.csv")


```

```{r}
i <- grep("CRISPLD2", resSig01$symbol)
resSig01[i,]
```

```{r}
plotCounts(dds, gene="ENSG00000103196", intgroup="dex")
```

```{r}
d <- plotCounts(dds, gene="ENSG00000103196", intgroup="dex", returnData=TRUE)
head(d)
```

```{r}
boxplot(count ~ dex , data=d)
```

```{r}
library(ggplot2)
ggplot(d, aes(dex, count)) + geom_boxplot(aes(fill=dex)) + scale_y_log10() + ggtitle("CRISPLD2")
```

```{r}
res$sig <- res$padj<0.05
table(res$sig)
sum(is.na(res$sig))
```


```{r}
library(ggplot2)

ggplot(as.data.frame(res), aes(log2FoldChange, -1*log10(pvalue), col=sig)) + 
    geom_point() + 
    ggtitle("Volcano plot")
```

